---
title: "Dominica: Process GFW positional data for voyages starting/ending in Dominica"
output: html_document
date: "2023-05-01"
---

```{r setup, include=FALSE}
# Packages
library(DBI)
library(bigrquery)
library(networkD3) # Make flow diagram
library(webshot) # Save static version of flow diagram

# Load common helper script
source(here::here("common.R"))

# Load reserve boundaries
load(here::here("results", "reserve_boundaries.Rdata"))

# Setup BQ connections
bq_auth_id <- "kmillage@ucsb.edu"

# Bigquery project for billing
bq_project <-  "emlab-gcp"

# Solution to bigrquery error (August 5, 2020) - see https://github.com/r-dbi/bigrquery/issues/395 for details
options(scipen = 20)
```

# Introduction

This script queries the GFW research messages tables to get more information for trips that start/end in Dominica. 

# Load reserve boundaries into Bigquery

Since we're interested in finding the last AIS position (for vessels arriving in Dominica) outside of the proposed reserve boundaries [or first in the case of vessels departing Dominica] we need to first upload the reserve boundaries we created to Bigquery as a spatial object.

```{r}
# Convert geometry list column to text that Bigquery will be able to read
reserve_boundaries_upload <- reserve_boundaries_sf %>%
  st_transform(crs = 4326) %>%
  mutate(geometry = st_as_text(geom, projjson = T)) %>%
  st_drop_geometry()

# Specify table fields
table_fields <- as_bq_fields(list(bq_field("method", "string"), bq_field("scenario", "integer"), bq_field("W_lon", "float"), bq_field("area_km2", "float"), bq_field("geometry", "geography")))

# Table to save output
reserve_boundaries_table <- bq_table(project = bq_project,
                                     table = "reserve_boundaries",
                                     dataset = "dominica")

# Upload
reserve_boundaries_table %>%
  bq_table_upload(fields = table_fields, values = reserve_boundaries_upload,
                  write_disposition = 'WRITE_TRUNCATE')
```
# Extract raw positional data for trips starting/ending in Dominica

Next we need to query the raw positional data for trips we identified as starting/ending in Dominica. This is an example query to pull those positions for 2022 - This took ~3 minutes and probably cost a few dollars. I think I need to figure out how to get this process filtered down.... 

```{r}
trip_info_sql <- 'WITH 

  ########################################
  # This subquery identifies good segments
  good_segments AS (
    SELECT
      seg_id
    FROM
      `world-fishing-827.pipe_production_v20201001.research_segs`
    WHERE good_seg
      AND NOT overlapping_and_short
    ),

  ########################################
  # This subquery extracts our trip information associated with voyages to/from ports in Dominica

  good_trip_info AS (
    SELECT
      *
    FROM
      `emlab-gcp.dominica.dominica_voyages_2017_2022`
  ),

  ########################################
  # This subquery gets all AIS messages 

  positions AS(
      SELECT 
        *
      FROM  
        `world-fishing-827.pipe_production_v20201001.research_messages`
      WHERE
        EXTRACT(YEAR FROM timestamp) > 2016 
        AND EXTRACT(YEAR FROM timestamp) < 2023
        AND _partitiontime BETWEEN "2022-01-01" AND "2022-12-31"
        AND seg_id IN (
          SELECT
              seg_id
            FROM
              good_segments)
  ),

  ########################################
  # This subquery filters the raw AIS positions  

  positions_keep AS(
    SELECT
    msgid,
    positions.ssvid,
    positions.seg_id,
    positions.timestamp,
    lat,
    lon,
    speed_knots,
    meters_to_prev,
    hours,
    source,
    distance_from_shore_m,
    trip_id
    FROM positions as positions
    INNER JOIN good_trip_info as trips 
      ON positions.ssvid = trips.ssvid
      AND positions.timestamp >= trips.start_timestamp 
      AND positions.timestamp <= trips.end_timestamp
  ),

  positions_inside AS(
    SELECT
      msgid,
      scenario
    FROM positions_keep as p
    JOIN `emlab-gcp.dominica.reserve_boundaries` AS reserve
    ON ST_WITHIN(st_geogpoint(p.lon, p.lat), reserve.geometry)
    WHERE lon IS NOT NULL and lat IS NOT NULL
  ),

  positions_matched AS(
    SELECT
      *,
      if(scenario is not null, TRUE, FALSE) as in_reserve
    FROM positions_keep
    LEFT JOIN positions_inside
      USING(msgid)
    ORDER BY
      trip_id,
      timestamp
  )

SELECT
  *
FROM
  positions_matched
ORDER BY
  trip_id,
  timestamp'
```


```{r query-port-index, eval=F}
# # Table to save output
# positions_2022_table <- bq_table(project = bq_project,
#                                  table = "all_relevant_positions_2022",
#                                  dataset = "dominica")
# 
# if(bq_table_exists(positions_2022_table)){
#   bq_table_delete(positions_2022_table)
# }else{
#   bq_table_create(positions_2022_table)
# }
# 
# # Run query - We have 25 named anchorages 
# bq_project_query(bq_project, 
#                  query = trip_info_sql,
#                  destination_table = positions_2022_table,
#                  use_legacy_sql = FALSE,
#                  allowLargeResults = TRUE)
```

Now let's try to summarize these positions by trip... 

```{r}
sql <- 'WITH 

  trip_totals AS (
    SELECT 
      trip_id,
      ssvid,
      COUNT(seg_id) as segments_total,
      SUM(hours) as hours_total,
      SUM(meters_to_prev) as dist_total,
      COUNT(msgid) as positions_total,
      MIN(timestamp) as start_timestamp,
      MAX(timestamp) as end_timestamp
    FROM
      `emlab-gcp.dominica.all_relevant_positions_2022`
    GROUP BY
      trip_id,
      ssvid
  ),

  outside_positions AS (
    SELECT
      *
    FROM
      `emlab-gcp.dominica.all_relevant_positions_2022`
    WHERE
      in_reserve = FALSE
  ),

   first_outside AS(
    SELECT
      trip_id,
      lat as first_lat_outside,
      lon as first_lon_outside,
      timestamp as first_timestamp_outside
    FROM (
      SELECT RANK() OVER (PARTITION BY trip_id ORDER BY timestamp) AS R,
        trip_id,
        lat,
        lon,
        timestamp
      FROM
      outside_positions
      ) AS ST
    WHERE ST.R = 1
  ),

  last_outside AS(
    SELECT
      trip_id,
      lat as last_lat_outside,
      lon as last_lon_outside,
      timestamp as last_timestamp_outside
    FROM (
      SELECT RANK() OVER (PARTITION BY trip_id ORDER BY timestamp) AS N,
        trip_id,
        lat,
        lon,
        timestamp
      FROM
      outside_positions
      ) AS ST
      WHERE ST.N = LAST_VALUE(ST.N)
  ),

  outside_totals AS (
    SELECT 
      trip_id,
      ssvid,
      SUM(hours) as hours_outside,
      SUM(meters_to_prev) as dist_outside,
      COUNT(msgid) as positions_outside
    FROM
      outside_positions
    GROUP BY
      trip_id,
      ssvid
  ),

  inside_positions AS (
    SELECT
      *
    FROM
      `emlab-gcp.dominica.all_relevant_positions_2022`
    WHERE
      in_reserve
  ),

  first_inside AS(
    SELECT
      trip_id,
      scenario,
      lat as first_lat_inside,
      lon as first_lon_inside,
      timestamp as first_timestamp_inside
    FROM (
      SELECT RANK() OVER (PARTITION BY scenario AND trip_id ORDER BY timestamp) AS P
      FROM
      inside_positions
      ) AS T
      WHERE T.P = 1
  ),

  last_inside AS(
    SELECT
      scenario,
      trip_id,
      lat as last_lat_inside,
      lon as last_lon_inside,
      timestamp as last_timestamp_inside
    FROM (
      SELECT RANK() OVER (PARTITION BY scenario AND trip_id ORDER BY timestamp) AS Y
      FROM
      inside_positions
      ) AS S
      WHERE S.Y = max(S.Y)
  ),

  inside_totals AS (
    SELECT 
      scenario,
      trip_id,
      ssvid,
      SUM(hours) as hours_inside,
      SUM(meters_to_prev) as dist_inside,
      COUNT(msgid) as positions_inside
    FROM
      inside_positions
    GROUP BY
      trip_id,
      ssvid,
      scenario
  )

SELECT 
  *
FROM
  trip_totals
JOIN outside_totals USING(trip_id, ssvid)
JOIN first_outside USING(trip_id, ssvid)
JOIN last_outside USING(trip_id, ssvid)
JOIN inside_totals USING(trip_id, ssvid)
JOIN first_inside USING(trip_id, ssvid, scenario)
JOIN last_inside USING(trip_id, ssvid, scenario)
'
```



