---
title: "Dominica: Process GFW positional data for voyages starting/ending in Dominica"
output: html_document
date: "2023-05-01"
---

```{r setup, include=FALSE}
# Packages
library(DBI)
library(bigrquery)
library(networkD3) # Make flow diagram
library(webshot) # Save static version of flow diagram

# Load common helper script
source(here::here("common.R"))

# Load reserve boundaries
load(here::here("results", "reserve_boundaries.Rdata"))

# Setup BQ connections
bq_auth_id <- "kmillage@ucsb.edu"

# Bigquery project for billing
bq_project <-  "emlab-gcp"

# Solution to bigrquery error (August 5, 2020) - see https://github.com/r-dbi/bigrquery/issues/395 for details
options(scipen = 20)
```

# Introduction

This script queries the GFW research messages tables to get more information for trips that start/end in Dominica. 

# Load reserve boundaries into Bigquery

Since we're interested in finding the last AIS position (for vessels arriving in Dominica) outside of the proposed reserve boundaries [or first in the case of vessels departing Dominica] we need to first upload the reserve boundaries we created to Bigquery. For some reason, the values got screwed up if I specified that the geometry column was of type "GEOGRAPHY" directly. So I'm first uploading the coordinates for the polygons as a string, and then converting it to the correct format once it's in Bigquery. 

```{r}
# Convert geometry list column to text that Bigquery will be able to read
reserve_boundaries_upload <- reserve_boundaries_sf %>%
  st_transform(crs = 4326) %>%
  mutate(geometry = st_as_text(geom, projjson = F)) %>%
  st_drop_geometry() %>%
  dplyr::select(scenario, geometry)

# Specify table fields
table_fields <- as_bq_fields(list(bq_field("scenario", "integer"), bq_field("geometry", "string")))

# Table to save output
reserve_boundaries_table <- bq_table(project = bq_project,
                                     table = "reserve_boundaries_raw",
                                     dataset = "dominica")

# Upload
reserve_boundaries_table %>%
  bq_table_upload(fields = table_fields, values = reserve_boundaries_upload,
                  write_disposition = 'WRITE_TRUNCATE')
```

```{r}
reserve_sql <- "
SELECT 
  scenario,
  ST_GEOGFROMTEXT(geometry) AS geom
FROM `emlab-gcp.dominica.reserve_boundaries_raw`"
```

```{r}
# Table to save output
reserve_table <- bq_table(project = bq_project,
                          table = "reserve_boundaries_fixed",
                          dataset = "dominica")

if(bq_table_exists(reserve_table)){
  bq_table_delete(reserve_table)
}else{
  bq_table_create(reserve_table)
}

# Run query
bq_project_query(bq_project,
                 query = reserve_sql,
                 destination_table = reserve_table,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE)
```

# Extract raw positional data for trips starting/ending in Dominica

Next we need to query the raw positional data for trips we identified as starting/ending in Dominica. This is an example query to pull the relevant positions for 2017 - 2022. This took ~3 minutes and probably cost $5. This shouldn't need to be rerun unless Tyler suggests a better way to match voyages to raw positions.

```{r}
relevant_positions_sql <- 'WITH 

  ########################################
  # This subquery identifies good segments
  good_segments AS (
    SELECT
      seg_id
    FROM
      `world-fishing-827.pipe_production_v20201001.research_segs`
    WHERE good_seg
      AND NOT overlapping_and_short
    ),

  ########################################
  # This subquery extracts our trip information associated with voyages to/from ports in Dominica

  good_trip_info AS (
    SELECT
      *
    FROM
      `emlab-gcp.dominica.dominica_voyages_2017_2022`
  ),

  ########################################
  # This subquery gets all AIS messages 

  positions AS(
      SELECT 
        *
      FROM  
        `world-fishing-827.pipe_production_v20201001.research_messages`
      WHERE
        EXTRACT(YEAR FROM timestamp) > 2016 
        AND EXTRACT(YEAR FROM timestamp) < 2023
        AND _partitiontime BETWEEN "2017-01-01" AND "2022-12-31"
        AND seg_id IN (
          SELECT
              seg_id
            FROM
              good_segments)
  ),

  ########################################
  # This subquery filters the raw AIS positions  

  positions_keep AS(
    SELECT
    msgid,
    positions.ssvid,
    positions.seg_id,
    positions.timestamp,
    lat,
    lon,
    speed_knots,
    meters_to_prev,
    hours,
    source,
    distance_from_shore_m,
    trip_id
    FROM positions as positions
    INNER JOIN good_trip_info as trips 
      ON positions.ssvid = trips.ssvid
      AND positions.timestamp >= trips.start_timestamp 
      AND positions.timestamp <= trips.end_timestamp
  )

SELECT
  *
FROM
  positions_keep
LEFT JOIN good_trip_info USING(trip_id)
ORDER BY
  trip_id,
  timestamp'
```

```{r query-positions, eval=F}
# # NOTE - DO NOT RERUN UNLESS ABSOLUTELY NECESSARY
# # Table to save output
# positions_table <- bq_table(project = bq_project,
#                             table = "all_relevant_positions_2017_2022",
#                             dataset = "dominica")
# 
# if(bq_table_exists(positions_table)){
#   bq_table_delete(positions_table)
# }else{
#   bq_table_create(positions_table)
# }
# 
# # Run query
# bq_project_query(bq_project, 
#                  query = relevant_positions_sql,
#                  destination_table = positions_table,
#                  use_legacy_sql = FALSE,
#                  allowLargeResults = TRUE)
```

# Match relevant positions to reserve boundaries

The next step is to determine which positions fall within our reserve boundaries. We're doing this in a separate step so we only have to query the 9 TB table once - this one is about 900 MB. 

```{r}
positions_spatial_sql <- 'WITH

  positions_inside AS(
    SELECT
      msgid,
      scenario,
    FROM `emlab-gcp.dominica.all_relevant_positions_2017_2022` as p
    JOIN `emlab-gcp.dominica.reserve_boundaries_fixed` AS reserve
    ON ST_WITHIN(st_geogpoint(p.lon, p.lat), reserve.geom)
    WHERE 
      lon IS NOT NULL AND lat IS NOT NULL

  ),
  
  positions_matched AS(
     SELECT
       *,
       IF(scenario IS NOT NULL, TRUE, FALSE) as in_reserve
     FROM `emlab-gcp.dominica.all_relevant_positions_2017_2022`
     LEFT JOIN positions_inside USING(msgid)
  )

SELECT
  *
FROM
  positions_matched
ORDER BY
  trip_id,
  timestamp'
```

```{r query-positions-spatial, eval=F}
# Table to save output
positions_spatial_table <- bq_table(project = bq_project,
                                    table = "all_positions_matched_to_reserve_2017_2022",
                                    dataset = "dominica")

if(bq_table_exists(positions_spatial_table)){
  bq_table_delete(positions_spatial_table)
}else{
  bq_table_create(positions_spatial_table)
}

# Run query
bq_project_query(bq_project,
                 query = positions_spatial_sql,
                 destination_table = positions_spatial_table,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE)
```

# Create a summary table for each trip

We want to know total distance, time, etc. for each trip, as well as the portions of those time(s) spent inside/outside the potential reserve areas. 

```{r}
trip_summary_sql <- "WITH 

  trip_totals AS (
    SELECT 
      trip_id,
      ssvid,
      COUNT(msgid) as messages_total,
      COUNT(seg_id) as segments_total,
      SUM(hours) as hours_total,
      SUM(meters_to_prev) as dist_total,
      MIN(timestamp) as timestamp_start,
      MAX(timestamp) as timestamp_end
    FROM
      `emlab-gcp.dominica.all_positions_matched_to_reserve_2017_2022`
    GROUP BY
      trip_id,
      ssvid
  ),

  inside_totals AS (
    SELECT 
      trip_id,
      ssvid,
      scenario,
      SUM(hours) as hours_in_reserve,
      SUM(meters_to_prev) as distance_in_reserve,
      MIN(timestamp) as first_timestamp_in_reserve,
      MAX(timestamp) as last_timestamp_in_reserve
    FROM
      `emlab-gcp.dominica.all_positions_matched_to_reserve_2017_2022`
    WHERE
      in_reserve
    GROUP BY
      trip_id,
      ssvid,
      scenario
  )

SELECT
  *
FROM
  trip_totals
LEFT JOIN inside_totals using (trip_id, ssvid)
ORDER BY
  trip_id,
  timestamp_start,
  scenario"
```

```{r}
# Table to save output
trip_summary_table <- bq_table(project = bq_project,
                                    table = "trip_summary_table_initial_2017_2022",
                                    dataset = "dominica")

if(bq_table_exists(trip_summary_table)){
  bq_table_delete(trip_summary_table)
}else{
  bq_table_create(trip_summary_table)
}

# Run query
bq_project_query(bq_project,
                 query = trip_summary_sql,
                 destination_table = trip_summary_table,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE)
```
## Trips into Dominica

Next, let's split our trips into those that are coming into Dominica, and those that are going out. Starting with those coming in, let's get some additional summary features for all of these trips.

```{r}
trips_coming_in_sql <- "WITH 

  voyages_into_dom AS (
    SELECT
      trip_id,
      ssvid,
      start_iso3,
      start_anchorage_id,
      start_anchorage_name,
      start_is_DMA,
      end_iso3,
      end_anchorage_id,
      end_anchorage_name,
      end_is_DMA
    FROM
      `emlab-gcp.dominica.dominica_voyages_2017_2022`
    WHERE
      start_is_DMA = FALSE 
      AND end_is_DMA
  ),

  vessel_info AS (
    SELECT DISTINCT
      CAST(mmsi as string) as ssvid,
      geartype as vessel_type
    FROM
      `world-fishing-827.vessel_database.all_vessels_20190102`
    WHERE
      geartype is not null
  ),
  
  all_relevant_positions AS (
    SELECT DISTINCT
      msgid,
      trip_id,
      ssvid,
      timestamp,
      lat,
      lon,
      speed_knots,
      meters_to_prev,
      hours
    FROM
      `emlab-gcp.dominica.all_positions_matched_to_reserve_2017_2022`
    WHERE
      trip_id IN (
          SELECT
            trip_id
          FROM
            voyages_into_dom)
  ),

  first_position AS(
      SELECT
        trip_id,
        lat as lat_start,
        lon as lon_start,
      FROM (
        SELECT RANK() OVER (PARTITION BY trip_id ORDER BY timestamp) AS i,
          trip_id,
          lat,
          lon
        FROM
          all_relevant_positions
      ) AS pos
      WHERE pos.i = 1
  ),

  last_position AS(
      SELECT
        trip_id,
        lat as lat_end,
        lon as lon_end,
      FROM (
        SELECT RANK() OVER (PARTITION BY trip_id ORDER BY timestamp desc) AS b,
          trip_id,
          lat,
          lon
        FROM
          all_relevant_positions
      ) AS pos
      WHERE pos.b = 1
  )

SELECT 
  trip_id,
  ssvid,
  vessel_type,
  start_iso3,
  start_anchorage_id,
  start_anchorage_name,
  start_is_DMA,
  end_iso3,
  end_anchorage_id,
  end_anchorage_name,
  end_is_DMA,
  messages_total,
  segments_total,
  hours_total,
  dist_total,
  timestamp_start,
  lat_start,
  lon_start,
  timestamp_end,
  lat_end,
  lon_end,
  scenario,
  hours_in_reserve,
  distance_in_reserve as dist_in_reserve,
  first_timestamp_in_reserve as timestamp_first_in_reserve,
  last_timestamp_in_reserve as timestamp_last_in_reserve
FROM
  voyages_into_dom
LEFT JOIN vessel_info USING (ssvid)
LEFT JOIN `emlab-gcp.dominica.trip_summary_table_initial_2017_2022` USING(trip_id, ssvid)
LEFT JOIN first_position USING (trip_id)
LEFT JOIN last_position USING (trip_id)
ORDER BY
  trip_id,
  scenario,
  timestamp_start"
```

```{r}
# Table to save output
trips_in_table <- bq_table(project = bq_project,
                           table = "trips_in_summary_table_raw",
                           dataset = "dominica")

if(bq_table_exists(trips_in_table)){
  bq_table_delete(trips_in_table)
}else{
  bq_table_create(trips_in_table)
}

# Run query
bq_project_query(bq_project,
                 query = trips_coming_in_sql,
                 destination_table = trips_in_table,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE)
```

Next we need to find the last position before each vessel entered the potential reserve for each scenario. I wasn't having any luck doing this for all scenarios at once, so we're going to loop over scenarios. 

```{r}
scenarios <- unique(reserve_boundaries_sf$scenario)

for(i in 1:length(scenarios)){

last_positions_sql <- paste0("WITH 

  trip_table AS (
    SELECT
      trip_id,
      scenario,
      timestamp_first_in_reserve
    FROM
      `emlab-gcp.dominica.trips_in_summary_table_raw`
    WHERE
      scenario = ", scenarios[i], "
  ),

  all_relevant_positions AS (
    SELECT DISTINCT
      msgid,
      trip_id,
      ssvid,
      timestamp,
      lat,
      lon,
      speed_knots,
      meters_to_prev,
      hours
    FROM
      `emlab-gcp.dominica.all_positions_matched_to_reserve_2017_2022`
    WHERE
      trip_id IN (
          SELECT
            trip_id
          FROM
            trip_table)
  ),

  points_before_inside AS(
      SELECT
        msgid,
        tp.trip_id,
        ssvid,
        timestamp,
        lat,
        lon,
        speed_knots,
        meters_to_prev,
        hours,
        tp.scenario
      FROM
        all_relevant_positions as ap
      JOIN trip_table as tp
        ON ap.trip_id = tp.trip_id
        AND ap.timestamp < tp.timestamp_first_in_reserve
  ),

  total_time_before_inside AS(
    SELECT
      trip_id,
      scenario,
      sum(hours) as hours_before_reserve,
      sum(meters_to_prev) as distance_before_reserve
    FROM
      points_before_inside
    GROUP BY
      trip_id,
      scenario
  ),

  last_point_before_inside AS(
      SELECT
        trip_id,
        scenario,
        timestamp as timestamp_last_before_reserve,
        lat as lat_last_before_reserve,
        lon as lon_last_before_reserve
      FROM (
        SELECT RANK() OVER (PARTITION BY trip_id, scenario ORDER BY timestamp desc) AS fi,
          trip_id,
          scenario,
          timestamp,
          lat,
          lon
        FROM
          points_before_inside
      ) AS ST
      WHERE ST.fi = 1
  )

SELECT 
  *
FROM
  trip_table
LEFT JOIN total_time_before_inside USING(trip_id, scenario)
LEFT JOIN last_point_before_inside USING(trip_id, scenario)
ORDER BY
  trip_id,
  scenario")

# Table to save output
trip_positions_table <- bq_table(project = bq_project,
                           table = paste0("trips_in_scenario_", scenarios[i]),
                           dataset = "dominica")

if(bq_table_exists(trip_positions_table)){
  bq_table_delete(trip_positions_table)
}else{
  bq_table_create(trip_positions_table)
}

# Run query
bq_project_query(bq_project,
                 query = last_positions_sql,
                 destination_table = trip_positions_table,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE)
}
```

And now let's combine all of this back to our summary table...

```{r}
trips_coming_in_final_sql <- "WITH

  all_trip_details AS (
    SELECT * FROM `emlab-gcp.dominica.trips_in_scenario_1`
    UNION ALL
    SELECT * FROM `emlab-gcp.dominica.trips_in_scenario_2`
    UNION ALL
    SELECT * FROM `emlab-gcp.dominica.trips_in_scenario_3`
  ),
  
  keep as (
    SELECT
      * EXCEPT(timestamp_first_in_reserve)
    FROM
      all_trip_details
  )

SELECT 
  *
FROM
  `emlab-gcp.dominica.trips_in_summary_table_raw`
LEFT JOIN keep  USING(trip_id, scenario)
ORDER BY
  trip_id,
  scenario"

# Table to save output
trips_in_table2 <- bq_table(project = bq_project,
                           table = "trips_in_summary_table_final",
                           dataset = "dominica")

if(bq_table_exists(trips_in_table2)){
  bq_table_delete(trips_in_table2)
}else{
  bq_table_create(trips_in_table2)
}

# Run query
trips_in <- bq_project_query(bq_project,
                 query = trips_coming_in_final_sql,
                 destination_table = trips_in_table2,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE) %>%
  bq_table_download()

# Save locally too
write_csv(trips_in, file.path(prj_cachalote_dir, "data", "processed", "trips_in_summary_table.csv"))
```

## Trips out of Dominica

Now to do the same with trips going out...

```{r}
trips_going_out_sql <- "WITH 

  voyages_out_of_dom AS (
    SELECT
      trip_id,
      ssvid,
      start_iso3,
      start_anchorage_id,
      start_anchorage_name,
      start_is_DMA,
      end_iso3,
      end_anchorage_id,
      end_anchorage_name,
      end_is_DMA
    FROM
      `emlab-gcp.dominica.dominica_voyages_2017_2022`
    WHERE
      start_is_DMA 
      AND end_is_DMA = FALSE
  ),

  vessel_info AS (
    SELECT DISTINCT
      CAST(mmsi as string) as ssvid,
      geartype as vessel_type
    FROM
      `world-fishing-827.vessel_database.all_vessels_20190102`
    WHERE
      geartype is not null
  ),
  
  all_relevant_positions AS (
    SELECT DISTINCT
      msgid,
      trip_id,
      ssvid,
      timestamp,
      lat,
      lon,
      speed_knots,
      meters_to_prev,
      hours
    FROM
      `emlab-gcp.dominica.all_positions_matched_to_reserve_2017_2022`
    WHERE
      trip_id IN (
          SELECT
            trip_id
          FROM
            voyages_out_of_dom)
  ),

  first_position AS(
      SELECT
        trip_id,
        lat as lat_start,
        lon as lon_start,
      FROM (
        SELECT RANK() OVER (PARTITION BY trip_id ORDER BY timestamp) AS i,
          trip_id,
          lat,
          lon
        FROM
          all_relevant_positions
      ) AS pos
      WHERE pos.i = 1
  ),

  last_position AS(
      SELECT
        trip_id,
        lat as lat_end,
        lon as lon_end,
      FROM (
        SELECT RANK() OVER (PARTITION BY trip_id ORDER BY timestamp desc) AS b,
          trip_id,
          lat,
          lon
        FROM
          all_relevant_positions
      ) AS pos
      WHERE pos.b = 1
  )

SELECT 
  trip_id,
  ssvid,
  vessel_type,
  start_iso3,
  start_anchorage_id,
  start_anchorage_name,
  start_is_DMA,
  end_iso3,
  end_anchorage_id,
  end_anchorage_name,
  end_is_DMA,
  messages_total,
  segments_total,
  hours_total,
  dist_total,
  timestamp_start,
  lat_start,
  lon_start,
  timestamp_end,
  lat_end,
  lon_end,
  scenario,
  hours_in_reserve,
  distance_in_reserve as dist_in_reserve,
  first_timestamp_in_reserve as timestamp_first_in_reserve,
  last_timestamp_in_reserve as timestamp_last_in_reserve
FROM
  voyages_out_of_dom
LEFT JOIN vessel_info USING (ssvid)
LEFT JOIN `emlab-gcp.dominica.trip_summary_table_initial_2017_2022` USING(trip_id, ssvid)
LEFT JOIN first_position USING (trip_id)
LEFT JOIN last_position USING (trip_id)
ORDER BY
  trip_id,
  scenario,
  timestamp_start"
```

```{r}
# Table to save output
trips_out_table <- bq_table(project = bq_project,
                           table = "trips_out_summary_table_raw",
                           dataset = "dominica")

if(bq_table_exists(trips_out_table)){
  bq_table_delete(trips_out_table)
}else{
  bq_table_create(trips_out_table)
}

# Run query
bq_project_query(bq_project,
                 query = trips_going_out_sql,
                 destination_table = trips_out_table,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE)
```

Next we need to find the last position before each vessel entered the reserve under each scenario. I wasn't having any luck doing this all at once, so we're going to loop over scenarios. 

```{r}
for(i in 1:length(scenarios)){

first_positions_sql <- paste0("WITH 

  trip_table AS (
    SELECT
      trip_id,
      scenario,
      timestamp_last_in_reserve
    FROM
      `emlab-gcp.dominica.trips_out_summary_table_raw`
    WHERE
      scenario = ", scenarios[i], "
  ),

  all_relevant_positions AS (
    SELECT DISTINCT
      msgid,
      trip_id,
      ssvid,
      timestamp,
      lat,
      lon,
      speed_knots,
      meters_to_prev,
      hours
    FROM
      `emlab-gcp.dominica.all_positions_matched_to_reserve_2017_2022`
    WHERE
      trip_id IN (
          SELECT
            trip_id
          FROM
            trip_table)
  ),

  points_after_inside AS(
      SELECT
        msgid,
        tp.trip_id,
        ssvid,
        timestamp,
        lat,
        lon,
        speed_knots,
        meters_to_prev,
        hours,
        tp.scenario
      FROM
        all_relevant_positions as ap
      JOIN trip_table as tp
        ON ap.trip_id = tp.trip_id
        AND ap.timestamp > tp.timestamp_last_in_reserve
  ),

  total_time_after_inside AS(
    SELECT
      trip_id,
      scenario,
      sum(hours) as hours_after_reserve,
      sum(meters_to_prev) as distance_after_reserve
    FROM
      points_after_inside
    GROUP BY
      trip_id,
      scenario
  ),

  first_point_after_inside AS(
      SELECT
        trip_id,
        scenario,
        timestamp as timestamp_first_after_reserve,
        lat as lat_first_after_reserve,
        lon as lon_first_after_reserve
      FROM (
        SELECT RANK() OVER (PARTITION BY trip_id, scenario ORDER BY timestamp) AS fi,
          trip_id,
          scenario,
          timestamp,
          lat,
          lon
        FROM
          points_after_inside
      ) AS ST
      WHERE ST.fi = 1
  )

SELECT 
  *
FROM
  trip_table
LEFT JOIN total_time_after_inside USING(trip_id, scenario)
LEFT JOIN first_point_after_inside USING(trip_id, scenario)
ORDER BY
  trip_id,
  scenario")

# Table to save output
trip_positions_table2 <- bq_table(project = bq_project,
                           table = paste0("trips_out_scenario_", scenarios[i]),
                           dataset = "dominica")

if(bq_table_exists(trip_positions_table2)){
  bq_table_delete(trip_positions_table2)
}else{
  bq_table_create(trip_positions_table2)
}

# Run query
bq_project_query(bq_project,
                 query = first_positions_sql,
                 destination_table = trip_positions_table2,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE)
}
```

```{r}
trips_coming_out_final_sql <- "WITH

  all_trip_details AS (
    SELECT * FROM `emlab-gcp.dominica.trips_out_scenario_1`
    UNION ALL
    SELECT * FROM `emlab-gcp.dominica.trips_out_scenario_2`
    UNION ALL
    SELECT * FROM `emlab-gcp.dominica.trips_out_scenario_3`
  ),
  
  keep as (
    SELECT
      * EXCEPT(timestamp_last_in_reserve)
    FROM
      all_trip_details
  )

SELECT 
  *
FROM
  `emlab-gcp.dominica.trips_out_summary_table_raw`
LEFT JOIN keep USING(trip_id, scenario)
ORDER BY
  trip_id,
  scenario"

# Table to save output
trips_out_table2 <- bq_table(project = bq_project,
                           table = "trips_out_summary_table_final",
                           dataset = "dominica")

if(bq_table_exists(trips_out_table2)){
  bq_table_delete(trips_out_table2)
}else{
  bq_table_create(trips_out_table2)
}

# Run query
trips_out <- bq_project_query(bq_project,
                 query = trips_coming_out_final_sql,
                 destination_table = trips_out_table2,
                 use_legacy_sql = FALSE,
                 allowLargeResults = TRUE) %>%
  bq_table_download()

# Save locally too
write_csv(trips_out, file.path(prj_cachalote_dir, "data", "processed", "trips_out_summary_table.csv"))
```

